Material Methods

### 2019-10-09 doc/MatMeth.Rmd
cleaned up lg in morning while mind fresh

Measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].

total Sc was measured [@wang_2019_sc]

Bivalent level measure were done [@peterson_2019]

@@

Statistical analysis was performed in R (v3.5.2)[@Rstudio].

The distributions of CO counts per cell was assessed for normailty (supplemental figure). Mouse mean MLH1 count was used to get around the bad discreet nature of count data. (normality was confirmed with X distribution, supplemental figure).

We choose to use a **mixed model framework** to interpert the patterns of variation. For interperting patterns of mean MLH1 count within and between house mouse subspecies.

1. write equation

$X \sim mouse average MLH1 ~ sex * subspecies + random(strain) + \xi $

In this model
mixed model, observations of mouse means of mlh1, all data with both male and female observations

The main effects and the interaction of subspecies and sex were coded as fixed.

Subspecies, sex and subsp*sex are fixed effects

Strain was coded as a random effect, since in this experimental design we consider it as a sample of natural genetic variation for each subspecies. We also included a term to allow for strains to have a specific sex effect.

This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect.

We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner)

From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.

We can also compare the size of the sex effects between (subspecies) and within (strain) mouse groups. (ie is the male or female strain effect larger? is there more variance due to random strain effect in male or females?)

I may have to do more research/reading on testing the random effects.
-specific statistical tests; (means, non-parametric kurts wallis t-test?) (correlation coefficient (function))
linear models lm() from X package.
-specific R packages
-non-parametric  (tests and measures for dealing with count data)
-hypothesis tested (size, DSB number)

Comparison/analysis of variation in MLH1 counts within and between groups was done within a general linear/ mix model frame work, using X package (cite).
With X formula Co count ~ blah + blah, with random strain specific sex effect.

(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)

>>>>

Hand measures of meiotic chromosomes generated by using ImageJ/Fiji (v1.52) [@fiji].

The total SC per cell and single bivalent measures were generated using the approaches discribed in [@wang_2019_sc] and [@peterson_2019] respectively.

@@

MLH1 counts per cell are good /decent proxies for genome wide recombination rates estimates (cite).

The distributions of CO counts per cell was assessed for normality (supplemental figure). Mouse mean MLH1 count was used to get around the bad discreet nature of count data. (normality was confirmed with X distribution, supplemental figure).

We chose to use a **mixed model framework** to interpret the patterns of variation, built using the lmer4 package (cite) in R (v3.5.2)[@Rstudio].

The use of wild derived inbred strains enabled us to design an utilize a mixed model for our analytical framework. We built our model to predict the strain average MLH1 count per cell. The main effects and their interaction of subspecies and sex were coded as fixed. While, strain (or genetic background) was coded as a random effect to reflect a random sample of natural genetic variation from each subspecies. A strains to have a specific sex effect was also included.

This model allows us to estimate and test the fixed effects (sex and subsp), and variation in mean MLH1 counts due to strain effect. We use the subspecies term as a way to quantify between group variation (or divergence) and the random strain effect to quantify within group variation (or polymorphism). (in a sex specific manner). From the mixed model the estimates of heterochiasmy (sexual dimorphism) are a combination of subspecies and strain specific sex effects.

We can also compare the size of the sex effects between (subspecies) and within (strain) mouse groups. (ie is the male or female strain effect larger? is there more variance due to random strain effect in male or females?)
(we used this frame work as proxy for 'polymorphism' and 'divergence' in MLH1 counts across house mouse and closely related species.)
I may have to do more research/reading on testing the random effects.

-specific R packages
-non-parametric  (tests and measures for dealing with count data)
-hypothesis tested (size, DSB number)

**outline**

1. framework for analyzing the MLH1 evolutionary patterns across mus species complex. (what is the main question we were tackling when designing this framework?)

2. write equation, describe terms (proxies for within and between polymorphism)
logic / choices for fixed and random effects

3. what are the 2 general / basic questions this framework can be applied to.

$X \sim mouse average MLH1 ~ sex + subspecies + random(strain) + \xi $

4. how can DMC1 measures be added to the mixed model

5. descriptions of mlh1 outside of framework (missing female data)
